---
title: "Week6_DataStructureTransform_Pirates"
author: "Anjali Rawat"
date: "February 16, 2019"
output: html_document
---

library(Rmarkdown)

# Q1. Combine the data into a single dataframe. Complete all the following exercises from the dataframe
```{r}
pirates=data.frame(Name=c("Astrid","Lea","Sarina","Remon","Letizia","Babice","Jonas","Wendy","Niveditha","Gioia"), Sex=c("F","F","F","M","F","F","M","F","F","F"),"Age"=c(30,25,25,29,22,22,35,19,32,21),Superhero=c("Batman","Superman","Batman","Spiderman","Batman","Antman","Batman","Superman","Maggott","Superman"),"Tattoos"=c(11,15,12,5,65,3,9,13,900,0))
View(pirates)
```

#Q2. What is the median age of the 10 pirates?
```{r}
median(pirates$Age)
```

# Q3.What was the mean age of female and male pirates separately?
```{r}
subset(pirates,Sex=="F")
        
piratesfemale=subset(pirates,Sex=="F")

MeanAgepiratesfemale=mean(piratesfemale$Age)

MeanAgepiratesfemale

piratesmale=subset(pirates,Sex=="M")
piratesmale

MeanAgepiratesmale=mean(piratesmale$Age)
MeanAgepiratesmale
```
#Q4. What was the most number of tattoos owned by a male pirate?
```{r}
MaxTattoospiratesales=max(piratesmale$Tattoos)
MaxTattoospiratesales
```
#Q5. What percent of pirates under the age of 32 were female?
```{r}
piratesunder32=subset(pirates,Age<32)
piratesunder32

femalepiratesunder32=subset(piratesunder32,Sex=="F")
femalepiratesunder32

percentpiratesunder32female=length(femalepiratesunder32$Sex)/length(piratesunder32$Sex)
percentpiratesunder32female
```

#Q6What percent of female pirates are under the age of 32?
```{r}
piratesfemale=subset(pirates,Sex=="F")
piratesfemale
piratesfemaleunder32=subset(piratesfemale,Age<32)
piratesfemaleunder32
length(piratesfemale$Sex)
length(piratesfemaleunder32$Sex)
percentpiratesfemaleunder32=length(piratesfemaleunder32$Sex)/length(piratesfemale$Sex)
percentpiratesfemaleunder32
```

#Q7Add a new column to the dataframe called tattoos.per.year which shows how many tattoos each pirate has for each year in their life.
```{r}
pirates$tattoos.per.year=pirates$Tattoos/pirates$Age

#Or using "with'

pirates$tattoos.per.year=with(Tattoos/Age,data=pirates)
pirates$tattoos.per.year
```


#Q8 Which pirate had the most number of tattoos per year?
```{r}
subset(pirates,pirates$tattoos.per.year==max((pirates$tattoos.per.year)))
```

#Q9What are the names of the female pirates whose favorite superhero is Superman?
```{r}
subset(pirates,pirates$Superhero=="Superman"& pirates$Sex=="F")
```


#Q10What was the median number of tattoos of pirates over the age of 20 whose favorite superhero is Spiderman?

```{r}
piratesover20=subset(pirates,pirates$Age>20)
piratesover20
        
piratesover20spiderman=subset(piratesover20,piratesover20$Superhero=="Spiderman")
piratesover20spiderman
``` 

